---
title: "Silk investment related analyses for *The web architecture, dynamics, and silk investment in the social spider, Stegodyphus sarasinorum*"
author: "Thejasvi"
date: 'Date of document initiation: Tuesday, March 29, 2016'
output:
  html_document: default
---

*This is one of three Rmd files accompanying the manuscript. The other files are  'beleyur_et_al_topology_analyses.Rmd' (analyses the coordination number and pore size data), 'weight_loss_analysis.Rmd' (analyses weight loss of individuals across colonies over the experiment)*

##### You are reading a version of the report compiled on `r date()`


Spiders were kept in 4 group sizes (1,5,10,25) and allowed to build webs on a 2x2feet wire frame. Using a custom-modified version of Angioquant, a freely available MATLAB toolbox, I extracted skeletonised forms of these webs over multiple spatial resolutions over the period of the experiment. 

### The dataset being analysed 'skeletondata,csv' has the following column headers:
### 1)colonyname 2)groupsize 3)Image name 4) Length of silk in grid 5) Number of junctions in grid
### 6)Serial number of grid (arranged column-wise from top-to-bottom) 7)Overal grid dimensions (eg. 32 means a 32x32 grid) 8) day, serial number of the reading from the start of experiment
  
--------------

```{r, echo=FALSE,cache=TRUE}
#dorig<-read.csv('C:/Users/Thejasvi/Google Drive/stego web/web-construction FINAL round - raw data/analysis-FINALround/analysis input-data/semifinal16thmay/skeletondata.csv')
setwd('./')
# also loading the multiplot function, which will be super useful as we go on:

#load the multiplot script that's sooo awesome in making panel graphs
source('../../analysis-scripts/smoothworkflow/multiplot.R')

# pixpm --> is the scaling factor from pixels --> metres.

# common address:

comad<-'../semifinal_16May/'


ps=1;
``` 


```{r, echo=FALSE,fig.width=12,fig.height=8}

## cleaning up the d$Length column because if increases and decreases ! 

#source(paste(comad,'Lengthclean.R',sep=''))

#TAIM<-system.time(procadjstlist<-prlladjstr()) # the processed list which has run thru Ladjstr

# ---- MAKING THIS THE FINAL DATASET FOR ALL ANALYSIS AND MANIPULATIONS !! 

#d<-do.call(rbind,procadjstlist) # rbinding all the lists into THE dataset

#-----loading it each time, instead of having to run all these commands! 

d<-read.csv(paste(comad,'finaldataset.csv',sep=''))
```

### Removing some spiders from the dataset

I only realised a couple of days ago that there are still some colonies which have 'extra' spiders in them. These extra spiders appear because 
the spiders somehow managed to move between webs despite our best efforts. These colonies will now be removed from the dataset. 

```{r}

excess_spiders <- read.csv('excess_spiders.csv')
colonies_to_remove <- subset(excess_spiders, excess_spiders_observed==TRUE)$colonyname


#load function which removes all rows in the colonynames list/vector
source('remove_colonies.R')

```


### A quick view of all the data together

###*Plot `r ps`: Length of silk vs time: Whole data :*


```{r,echo=FALSE}
# now making the plots of the 'corrected' data 
library(ggplot2)

wd<-subset(d,d$GridSize==1)

# Pending any further corrections - remove the excess spider colonies (as of 2020-09-02 14:02 CET)
wd <- remove_colonies(wd, colonies_to_remove,'colonyname')

q025<-function(v){return(quantile(v,0.025,na.rm=T))} # gives output of 2.75%ile
q975<-function(v){return(quantile(v,0.975,na.rm=T))} # gives output of 97.5%iles

#   a plot to see the numbers in metres
# 1536 pixels== 24 inches ==> 60.96 cm--> 1536/60.96 --> 
pixpm<-1536/.6096     # number of pixels per m 

wd$Lengthm<-wd$Length/pixpm

p1mtrs<-ggplot(wd,aes(as.factor(day),Length/pixpm,color=as.factor(groupsize),group=colonyname)) 
p11mtrs<-p1mtrs+geom_point()+geom_line(linetype='solid',size=0.75)+ylab('Length of silk (metres) ') + xlab('Days');

p2mtrs<-ggplot(wd,aes(day,Length/pixpm,color=as.factor(groupsize),group=groupsize))  + stat_summary(fun.y = median, fun.ymin = q025,fun.ymax =q975 ,size=1.5) +ylab('')+xlab('Days')+scale_x_discrete(breaks=c(1:10),labels=c(1:10));



# figure 3 : plot with standard errors

# now need to make a new plot with days on x , mean + se on y axis for each group size :
mnwd<-aggregate(Lengthm ~ day*groupsize, data = wd, mean);colnames(mnwd)[3]<-'meanL'
sdwd<-aggregate(Lengthm ~ day*groupsize, data = wd, sd);colnames(sdwd)[3]<-'sdL'
# merged and ordered
mgwd<-merge(mnwd,sdwd); mgwd<-mgwd[order(mgwd[,1],mgwd[,2]),]

#
pd <- position_dodge(width = 0.8)
mgwd1<-subset(mgwd,day<10)
mgwd7<-subset(mgwd,day<8)


mgplt<- ggplot(mgwd1,aes(x=as.factor(day),y=meanL,shape=as.factor(groupsize),group=as.factor(groupsize))) +geom_pointrange(aes(ymin=meanL-sdL, ymax=meanL+sdL,width=0.9),size=1.3,position=pd) +scale_shape_manual(values=c(15,16,17,18))+ xlab('Days') + ylab('Length of silk (metres)') +geom_line(lwd=1,position=pd)+labs(shape = "Group Size")+ theme(text = element_text(size=15))  ; mgplt
ggsave('total_silk_length_all_10days.png')



# final figures as of 2020-05 limited to 7 days

mgplt7<- ggplot(mgwd7,aes(x=as.factor(day),y=meanL,shape=as.factor(groupsize),group=as.factor(groupsize))) +geom_pointrange(aes(ymin=meanL-sdL, ymax=meanL+sdL,width=0.9),size=1.3,position=pd) +scale_shape_manual(values=c(15,16,17,18))+ xlab('Days') + ylab('Length of silk (metres)') +geom_line(lwd=1,position=pd)+labs(shape = "Group Size")+ theme(text = element_text(size=15))  ; mgplt7

ggsave('total_silk_length_all_7days.png')




ps<-ps+1;


multiplot(p11mtrs,p2mtrs,cols=2)


```

#####*The lines show the 95% interval of the data from the median (2.75% and 97.25%)

#### There seems to be a decent amount of overlap in silk investment across group sizes - but the 25 spider group seems to be quite distinct. This points to quite a bit of colony level variation. 

### *Plot `r ps`*: How does the amount of silk scale with groupsize across time ?

```{r,width=12,fig.height=8}

#   SHOWS THE MEDIAN VALUE AT TIME WITH GROUP-SIZE ON X-AXIS
daywiseplots<-ggplot(wd,aes(x=as.factor(groupsize),y=Length/pixpm,color=as.factor(day),group=as.factor(day)))  + stat_summary(fun.y = median,fun.ymin=q025,fun.ymax=q975,size=1.5,geom='line') +ylab('')+xlab('groupsize')+scale_x_discrete(breaks=c(1,5,10,25),labels=c(1,5,10,25))

daywiseplots

ps<-ps+1
```


```{r}

M1p<-ggplot(wd,aes(x=day,y=log10(Length/pixpm+0.001),color=as.factor(groupsize),group=groupsize))+geom_smooth(lwd=2)+ylim(0,2.5)+ylab('Log10 length of silk on web (metres)')+theme(legend.position='none')

 M2p<-ggplot(wd,aes(x=day,y=Length/pixpm,color=as.factor(groupsize),group=groupsize))+geom_smooth(lwd=2)+ylab('Length of silk on web (metres)')+theme(legend.position=c(0.25,0.95))

multiplot(M2p,M1p,cols=2)
```




### *Plot `r ps `* : $\Delta\$ silk investment over time:

```{r,echo=FALSE,warning=FALSE, fig.width=15, fig.height=8,message=FALSE}
#assigning a new column:

# OMG the 'by'  function is soooo neat:
wddif<-by(wd$Length,wd$colonyname,diff) # apply the function for every colonyname and get the output 

#now to join up all these vectors:
library(qpcR);
library(reshape2);
fun<-qpcR:::cbind.na
dlength<-do.call(fun,wddif) # a columnwise data.frame with change in lengths
#convert this 'wide' formate to a 'long' format using 'melt' in reshape2
dlenth<-melt(dlength);names(dlenth)<-c('day','colonyname','dLength');

# removes the last alphabet from the colonyname
endremv<-function(nem){ if (is.na(nem)==FALSE) {as.numeric(substr(nem, 1, nchar(nem)-1))} else if (is.na(nem)==TRUE){return(NA)} } 

dlenth$groupsize<-as.numeric(lapply(as.character(dlenth$colonyname),endremv)) # and we have our final dataset:

dg1<-ggplot(dlenth,aes(x=day,y=dLength/pixpm,group=colonyname,colour=as.factor(groupsize)))+geom_point()+geom_line(linetype='solid',size=0.75)+ylab('Growth in silk investment (metres)') + xlab('Days')+scale_x_discrete(breaks=c(1:9),labels=c(1:9))+theme(legend.position='none')

dg2<-ggplot(dlenth,aes(x=day,y=dLength/pixpm,color=as.factor(groupsize),group=groupsize))  + stat_summary(fun.y = median, fun.ymin = q025,fun.ymax =q975 ,size=1.5,position=position_dodge(width=0.65)) +ylab('')+xlab('Days')+scale_x_discrete(breaks=c(1:9),labels=c(1:9))+theme(legend.position=c(0.85,0.85));

multiplot(dg1,dg2,cols=2)

#dg2+ylab('Daily web silk investment (m)')+theme(legend.title=element_blank())
dg3<-ggplot(dlenth,aes(x=day,y=dLength/pixpm,color=as.factor(groupsize),group=groupsize))+geom_point()  + stat_smooth(lwd=2,level=0.95) +ylab('')+xlab('Days')+scale_x_discrete(breaks=c(1:9),labels=c(1:9))+theme(legend.position=c(0.85,0.85))+theme(legend.title=element_blank())+ylab('Daily silk investment (metres)');dg3




ps<-ps+1
```


### *Plot `r ps `* : Per Capita $\Delta\$ silk investment :
```{r,echo=FALSE,warning=FALSE,fig.width=12,fig.height=8}

# divide the dLength/groupsize - make a new column of it :

dlenth$percap<-(dlenth$dLength/dlenth$groupsize)/pixpm # percapita in metres

#ggplot(dlenth,aes(x=day,y=percap/pixpm,color=as.factor(groupsize),group=groupsize)) 

pcp1<-ggplot(dlenth,aes(x=day,y=percap/pixpm,color=as.factor(groupsize),group=as.factor(groupsize)))  + stat_summary(fun.y = median, fun.ymin = q025,fun.ymax =q975 ,size=1.5,position=position_dodge(width=0.65)) +ylab('Per-capita silk investment (m)')+xlab('Days')+scale_x_discrete(breaks=c(1:9),labels=c(1:9))+theme(legend.position=c(0.75,0.85))+theme(legend.title=element_blank());pcp1

pcp3<-ggplot(dlenth,aes(x=day,y=percap/pixpm))+geom_point(size=4)+stat_smooth()+ylab('Per-capita silk investment (m)') + xlab('Days')+scale_x_discrete(breaks=c(1:9),labels=c(1:9))+theme(legend.position=c(0.75,0.85))+theme(legend.title=element_blank());pcp3


# Parameters from figure 3 - for consistency
#mgplt<- ggplot(mgwd1,aes(x=as.factor(day),y=meanL,group=as.factor(groupsize))) +geom_pointrange(aes(ymin=meanL-sdL, ymax=meanL+sdL,shape=as.factor(mgwd1$groupsize),width=0.9),size=1.3,position=pd) + xlab('Days') + ylab('Length of silk (metres)') +geom_line(lwd=1,position=pd)+labs(shape = "Group Size")+ theme(text = element_text(size=15)) ; mgplt

# This is a broken piece of code that isn't part of  the manuscript - and so am not bothering to fix right now. 

# pcp2<-ggplot(dlenth,aes(x=as.factor(day),y=percap,group=groupsize,shape=as.factor(dlenth$groupsize),linetype=as.factor(dlenth$groupsize))) + #geom_point(size=3)+geom_smooth(size=2,alpha=0.2)+ylab('Per-capita silk investment (m)') + xlab('Days')+scale_x_discrete(breaks=c(1:9),labels=c(1:9))+labs(shape = "Group Size (Raw data)")+ #theme(text = element_text(size=20))+scale_linetype_manual(values=c('solid','dashed','dotdash', "dotted"),name="Group size (LOESS)", labels = c("1", "5", "10",'25') #)+theme(legend.key.width = unit(2.5,"cm"));pcp2
# +scale_linetype_discrete()

# now to make the pcp2 plot to show the averaged per-capita silk with se:

# figure 4 : 

# now need to make a new plot with days on x , mean + se on y axis for each group size :
mnpc<-aggregate(percap ~ day*groupsize, data = dlenth, mean);colnames(mnpc)[3]<-'meanPC'
sdpc<-aggregate(percap ~ day*groupsize, data = dlenth, sd);colnames(sdpc)[3]<-'sdPC'
# merged and ordered
mgpc<-merge(mnpc,sdpc); mgpc<-mgpc[order(mgpc[,1],mgpc[,2]),]

#### figure  4 
pd <- position_dodge(width = 0.8)
mgpc1<-subset(mgpc,day<9)
mgpltpc<- ggplot(mgpc1,aes(x=as.factor(day),y=meanPC,shape=as.factor(groupsize),group=as.factor(groupsize))) +geom_pointrange(aes(ymin=meanPC-sdPC, ymax=meanPC+sdPC,shape=as.factor(mgpc1$groupsize),width=0.9),size=1.3,position=pd) + xlab('Days') + ylab('Per-capita silk investment (metres)')+scale_shape_manual(values=c(15,16,17,18))+geom_line(lwd=1,position=pd)+labs(shape = "Group Size")+ theme(text = element_text(size=20))  ; mgpltpc
ggsave('per_capita_all_10days.png')



### figure 4 -- within 7 days
mgpc7<-subset(mgpc,day<8)

mgpltp7<- ggplot(mgpc7,aes(x=as.factor(day),y=meanPC,shape=as.factor(groupsize),group=as.factor(groupsize))) +geom_pointrange(aes(ymin=meanPC-sdPC, ymax=meanPC+sdPC,shape=as.factor(mgpc7$groupsize),width=0.9),size=1.3,position=pd) + xlab('Days') + ylab('Per-capita silk investment (metres)')+scale_shape_manual(values=c(15,16,17,18))+geom_line(lwd=1,position=pd)+labs(shape = "Group Size")+ theme(text = element_text(size=20))  ; mgpltp7
ggsave('per_capita_all_7days.png')


```


#### Is the per-capita silk investment uniform across time and groupsize ? 
#### the results suggests not - which means that the spiders are adding the same amount of silk across time and groupsize. P.S. The 'warning' that appears does not affect the reliability of the test results - pers comm with package author)

### Running the nparLD analysis
The nparLD analysis must be run with an older version of R (here I'm using 2.15.3). To do so, 
I'm using the Renv (https://github.com/viking/Renv) and R-build https://github.com/viking/R-build
projects. For a comparison of how to do this interactively - please see the topology analysis file. 

```{r}
# this activates Renv. Make sure the default R version is set to 2.15.3
# this is a Linux specific way to do it 
save(dlenth, file='dlenth.Rdata')
save(wd, file='wd.Rdata')

system('bash -c "source ~/.bash_profile"')
system('bash -c "Rscript silk_nparLD.R"')

# also - don't worry about the 'warning :covariance matrix is singular' - 
# pers comm from 1st author off the package that it doesn't affect the results of the test

load('percapita_nparld_results.Rdata')

print(res$ANOVA.test)


# also load the overall silk nparlD results 
load('overallsilk_nparLD_results.Rdata')
print(overall_silk$ANOVA.test)

```

### Other general descriptive statistics in the paper

#### Number of replicates per group size

```{r, echo=TRUE}
wd1<-subset(wd, day==1)
table(wd1$groupsize)
```

### The R packages and R version required to run this analysis 

```{r}
# code from 
# https://www.r-bloggers.com/list-of-user-installed-r-packages-and-their-versions/

ip <- as.data.frame(installed.packages()[,c(1,3:4)])
rownames(ip) <- NULL
ip <- ip[is.na(ip$Priority),1:2,drop=FALSE]
print(ip, row.names=FALSE)

## The R version used to run this analysis 
print(R.version)
```

