---
title: "silk_density_with_radial_distance"
author: "Thejasvi"
date: "30 August 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This script will present plots of how silk density varies with radial distance from the retreat
## Let's just take a look at one such silk density heatmap below:

```{r , echo=FALSE,}

# use the previously existing dataset of smoothened grid values 

sm_g64_path = 'D:\\stego web\\web-construction FINAL round - raw data\\analysis-FINALround\\analysis-results\\semifinal_16May\\64GridSmoothenedarray.RDS'

d<-readRDS(sm_g64_path)

# just take an example dataset
d_eg <-d$'10E'
d_eg1 <- d_eg[,,3] # take the 4th day 
heatmap(d_eg1,Rowv=NA,Colv=NA)

```

## Now let's setup 3 radial arcs of one gridbox width from the bottomr right tip of the retreat:


```{r make radial lines, echo=FALSE}

tip_index = c(5,5) # for the large retreats the retreat was 2x2 inches in size
bottom_tip = c(64,64)
diagonal_dist = round( dist( rbind(tip_index,bottom_tip) ) )

# we take silk density sections at 1/4th, 1/2 and 3/4th of this maximum possible diagonal distance

arc_sections = round ( diagonal_dist * c(1/4,2/4,3/4)) # the radial distances at which the arcs 
                                                       # will be measured at 

xy_indxs = expand.grid(1:64,1:64)

#calculate all grids that are within the arc_sections 


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
