---
title: "Weight loss analysis for *The web architecture, dynamics, and silk investment in the social spider, Stegodyphus sarasinorum*"
author: "Thejasvi Beleyur"
output:
  html_document: default
---

*This is one of three Rmd files accompanying the manuscript. The other files are called 'beleyur_et_al_silk_analyses.Rmd' (analyses the silk length on the web and the per-capita silk investment.') and 'beleyur_et_al_topology_analyses.Rmd' (analyses the coordination number and pore size data)*

This notebook analyses the weight loss of individuals across the experiment to explore how individuals lost weight as a result of dessication + silk investment with group size. Here, the outcome variable is the weight loss and the predictor variables used will be:
1. weighing day (#days after experiment start)
1. body condition ($\frac{starting\:body\:weight}{cephalothorax\:width}$)
1. group-size

Colony name will also be used as a random effect, to account for colony-specific differences in silk-investment and dessication of spiders. 

In general, the approach used here is to inspect model outputs from the default frequentist output using the  ```lmer``` function, followed by Bayesian parameter estimation using non-informative priors with the ```sim``` function. 

```{r, loading, results='hide', warning=F, message=F}
library(arm)
library(coda)
library(dplyr)
library(gplots)
library(lme4)
set.seed(82319)
wtadrs<-'../../analysis input-data/finalwtsnsizes.csv' # address of the weights and sizes csvfile
wts_raw<-read.csv(wtadrs)

# 2020-09-02 : realised I had left some of the colonies in the dataset with excess spiders (caused by
# spiders ballooning their way from one web to the other) - and these need to be removed. 

excess_spiders <- read.csv('excess_spiders.csv')
colonies_to_remove <- subset(excess_spiders, excess_spiders_observed==TRUE)$colonyname

#load function which removes all rows in the colonynames list/vector
source('remove_colonies.R')

wts <- remove_colonies(wts_raw, as.vector(colonies_to_remove),'colonyname')

wts<-subset(wts,wts$groupsize!='NA') # cleaning up the data frame once more to avoid an extra 'NA' group ! 
wts$wtloss<-wts$b4wt-wts$aftrwt

  
hqt<-round(quantile(wts$wtloss,0.975,na.rm=T),4)
lqt<-round(quantile(wts$wtloss,0.025,na.rm=T),4)
medn<-round(quantile(wts$wtloss,0.5,na.rm=T),4)


# now converting all the wts into milligrams with 1 decimal point accuracy - this is just to ease interpretation

wts$wtlossmg<-round(wts$wtloss*1000,1)
wts$b4wtmg<-round(wts$b4wt*1000,1)
wts$aftrwtmg<-round(wts$aftrwt*1000,1)
wts$wtlossmg<-round(wts$wtloss*1000,1)


# how much of missing data -(for Weightloss)  is there in this dataset ?
rws<-dim(wts)[1]
mssng<-sum(is.na(wts$wtloss))
pstagemiss<-mssng/rws

#are weight loss and spider size connected ? - let's check this out ! 
plot(wts$cephwid,wts$wtloss*1000,pch=20,col='red')
cor.test(wts$cephwid,wts$wtloss,method='spearman')

# generate body condition 
wts$body.cond <- wts$b4wtmg/wts$cephwid

wts$fact.gsize <- as.factor(wts$groupsize)
# standardise cephalothorax width
wts$std.cephwid <- (mean(wts$cephwid,na.rm=T)-wts$cephwid)/sd(wts$cephwid, na.rm=T)
wts$std.b4wtmg <- (mean(wts$b4wtmg,na.rm=T)-wts$b4wtmg)/sd(wts$b4wtmg, na.rm=T)
wts$std.wtlossmg <- (mean(wts$wtlossmg,na.rm=T)-wts$wtlossmg)/sd(wts$wtlossmg, na.rm=T)
wts$std.body.cond <- (mean(wts$body.cond,na.rm=T)-wts$body.cond)/sd(wts$body.cond, na.rm=T)



#dlenth <- subset(dlenth)
get.hpdintervals <- function(x,probrange=0.93)
{HPDinterval(as.mcmc(x),prob=probrange)}
get.med.hpdintervals <- function(x,probrange=0.93)
          {
          hpd.interval<-get.hpdintervals(x,probrange)
          med <- median(x)
          output <- c(hpd.interval[1],med,hpd.interval[2])
}
get.prob.lesser.value <- function(x,value){mean(x<value)}

wts.nona <- wts[complete.cases(select(wts,b4wtmg,wtday,fact.gsize,cephwid,std.cephwid)),]
```

Now, let's build the models.

```{r,echo=TRUE,warning=F}

# both of these results suggest that there's an influence of previous weight, and size on silk investment
# maybe it's best to do some kind of linear model first to figure this out :

M0<-lm(wtlossmg~b4wtmg+wtday+fact.gsize+std.cephwid,data=wts.nona,
                  na.action="na.exclude")
res.M0 <- residuals(M0)
summary(M0)
frame()
par(mfrow=c(3,3))
plot(M0)
scatter.smooth(wts.nona$wtday,res.M0)
scatter.smooth(wts.nona$cephwid,res.M0)
boxplot(res.M0~as.factor(wts.nona$colonyname))
boxplot(res.M0~wts.nona$fact.gsize)


```

The residual inspection seems to indicate an okay fit, with 2 exceptions : colony-level variation and a non-linear pattern in the day factor. However, this non-linear pattern may also be affected by the fact that different colonies were sampled after different number of days. So let's just move straight to the linear mixed model now. 



```{r, echo=TRUE,,warning=F}
ml0<-lmer(std.wtlossmg~std.body.cond+wtday+fact.gsize+(1|colonyname),data=wts.nona, na.action="na.exclude")
summary(ml0)

frame()
par(mfrow=c(3,3))
# residuals vs group size
scatter.smooth(as.factor(wts.nona$groupsize), resid(ml0))
# residuals vs day 
scatter.smooth(wts.nona$wtday,resid(ml0))
# sqrt(abs(residuals))
scatter.smooth(wts.nona$wtday, sqrt(abs(resid(ml0))))
# sqrt resids - groupsize
scatter.smooth(as.factor(wts.nona$groupsize), sqrt(abs(resid(ml0))))
# residuals vs body condition
scatter.smooth(wts.nona$std.body.cond,resid(ml0));abline(0,0)
# residuals vs colonyname 
boxplot(resid(ml0)~as.factor(wts.nona$colonyname));abline(0,0)
# residuals vs b4wt
scatter.smooth(resid(ml0)~wts.nona$std.b4wtmg);abline(0,0)
# qqplot model residuals
qqnorm(resid(ml0));qqline(resid(ml0))
# qqplot random effects
qqnorm(ranef(ml0)$colonyname$'(Intercept)');
qqline(ranef(ml0)$colonyname$'(Intercept)')

```

The residual inspection here shows a much better fit. The weird bump on day 8 in residuals is gone, this must be because of colony-level variation. It is important to note that cephalothorax width has been standardised here. This means an increase of one std. in cephwidth will have a corresponding change in weightloss in mg. 


```{r ml0.sim, warning=F}

ml0.sim <- sim(ml0, n.sim=5000)
# extract th 93% compat. interval of all coefficients
ml0.hpd.coefs <- apply(ml0.sim@fixef, 2, get.med.hpdintervals)
print(ml0.hpd.coefs)

gpsize.coefs <- ml0.sim@fixef[,1]+ml0.sim@fixef[,4:6]
gpsize.hpd.coefs <- apply(gpsize.coefs, 2, get.med.hpdintervals)
print(gpsize.hpd.coefs)

gsize <- seq(4,6)
non.gsize <- c(2,3)
frame()
par(mfrow=c(1,2))
plotCI(seq(1,2),y=ml0.hpd.coefs[2,non.gsize],li=ml0.hpd.coefs[1,non.gsize],ui=ml0.hpd.coefs[3,non.gsize], main='Fixed effects',ylab = 'Coefficients', xlab=' ',axes=F,las=2,ylim=c(-1,1))
axis(1, at=1:2, labels=colnames(ml0.hpd.coefs)[non.gsize],las=2);axis(2, at=seq(-4,4,), labels=seq(-4,4,))
plotCI(seq(1,3),y=gpsize.hpd.coefs[2,],li=gpsize.hpd.coefs[1,],ui=gpsize.hpd.coefs[3,], main='Fixed effects',ylab = 'Coefficients', xlab=' ',axes=F,las=2,ylim=c(-4,4))
axis(1, at=1:3, labels=colnames(ml0.hpd.coefs)[gsize],las=2);axis(2, at=seq(-4,4,), labels=seq(-4,4,))

```

The model informs us that individuals with better body-condition tend to lose more weight than those with lower body condition. Day of weighing seems to have an overall positive effect on weight loss, which makes sense. The $P(wtday<0)=$ `r round(mean(ml0.sim@fixef[,3]<0),2)`. Remember that the outcome is a standardised variable, and the coefficients are to be interpreted as one day increase in time corresponds to X sd increase in weight loss. So, for between 6-10 days of time, this corresponds to a median of `r round(ml0.hpd.coefs[2,3]*c(6),1)`-`r round(ml0.hpd.coefs[2,3]*c(10),1)` increase in weight loss std. deviations. This comparison between time and body condition tells us that body condition is a much stronger effect than time. 

In general, there is a (weakly) indicative trend that larger groupsizes lead to lower weight loss, though its is somewhat weak. We see the $P(Coef.(group\:sizes)<0)=$ `r round(apply(ml0.sim@fixef[,4:6],2,get.prob.lesser.value,0),2)`. The decrease in weight loss with increasing group size is most strongly seen in group-size 25 of course, where the $P(Coef.(group\:size))<0$ is ~ `r round(get.prob.lesser.value(ml0.sim@fixef[,6],0),1)`. For instance, spiders in group-size 25 are expected to show a ~ median of `r round(ml0.hpd.coefs[2,6],1)` std.deviations decrease in weight loss, while other group sizes show somewhat lower effects. If we would like to compare the effects of group size and body condition on weight loss, we  need to compare the coefficients:

```{r groupsize.vs.bodycond}

p.gsize.greater.bodycond <- c(0,0,0)
for (i in seq(1,3)){
  p.gsize.greater.bodycond[i]<-mean(gpsize.coefs[,i]>ml0.sim@fixef[,2])}


print(round(p.gsize.greater.bodycond,1))

```
Here we see that in general, the effect of body condition is greater than that of group-size only ~30% of the time. What about the relative coefficients? 

```{r rel.groupsize.bodycond}
rel.effects.gsize <- abs(gpsize.coefs)/abs(ml0.sim@fixef[,2])
print(apply(rel.effects.gsize,2,get.med.hpdintervals,probrange=0.5))
# calculate prob>1 
p.gsize.greater.bodycond <- apply(rel.effects.gsize,2,function(x){mean(x>1)})
print(round(p.gsize.greater.bodycond,2))
```
The contribution of group-size is in general more than body-condition : $P(\frac{|Coef(groupsize)|}{|Coef(body\:condition)|}>1=$ `r round(p.gsize.greater.bodycond,1)` for groupsizes 5,10,25.



Let's test for the effect of batch - this tries to account for weather conditions (sun/rain) that may have affected dessication indirectly by affecting the temperature of the experimental room. 

```{r, echo=TRUE}
lastchar<-function(V){substr(V,nchar(V),nchar(V))}
# different batches were run at different points of time.
wts.nona$batchid<-as.factor(lastchar(as.character(wts.nona$colonyname)))

ml0b<-lmer(std.wtlossmg~std.body.cond+wtday+fact.gsize+(1|batchid)+(1|colonyname)-1,data=wts.nona, na.action="na.exclude")
summary(ml0b)
```

...and so - batch ID doesn't seem to have an effect on the weight loss.



### Other descriptive data

```{r,echo=TRUE}

# number of spiders weighed
numwd<-sum(is.na(wts$b4wt)==F)
# number of spiders re-weighed
numrewd<-sum(is.na(wts$aftrwt)==F)

# percentage of spiders re-weighed: ~80%
pctgrwd<-numrewd/numwd

# -- average weight of spiders b4 :
mean(wts$b4wtmg,na.rm=T)
```



#### Cephalothorax widths

```{r, echo=TRUE}
cephalothorax_width = wts$cephwid
mean_cephwid <- mean(cephalothorax_width, na.rm=TRUE)
std_cephwid <- sd(cephalothorax_width, na.rm=TRUE)
sample_size = length(cephalothorax_width) - length(sum(is.na(cephalothorax_width)))
sprintf('Mean cephalothorax width +/- sd: %f +/- %f mm, N=%d', mean_cephwid, std_cephwid, sample_size)
```


```{r, echo=TRUE}
mean_startwt <- mean(wts$b4wt, na.rm=TRUE)
std_startwt <- sd(wts$b4wt, na.rm=TRUE)
sample_size = length(wts$b4wt) - length(sum(is.na(wts$b4wt)))

sprintf('Mean cephalothorax width +/- sd: %f +/- %f mg, N=%d', mean_startwt*1000, std_startwt*1000, sample_size)

```

## Working environment 
This information has a lot of the information to replicate the computational setup required to replicate the results in this analysis. 

```{r workenv}
print(sessionInfo())


```


