---
title: "Weight loss analysis for *The web architecture, dynamics, and silk investment in the social spider, Stegodyphus sarasinorum*"
author: "Thejasvi"
output:
  html_document: default
---

*This is one of three Rmd files accompanying the manuscript. The other files are called 'beleyur_et_al_silk_analyses.Rmd' (analyses the silk length on the web and the per-capita silk investment.') and 'beleyur_et_al_topology_analyses.Rmd' (analyses the coordination number and pore size data)*

This notebook analyses the weight loss of individuals across the experiment to explore how individuals lost weight as a result of dessication + silk investment with group size.

```{r, echo=TRUE}
wtadrs<-'../../analysis input-data/finalwtsnsizes.csv' # address of the weights and sizes csvfile
wts_raw<-read.csv(wtadrs)

# 2020-09-02 : realised I had left some of the colonies in the dataset with excess spiders (caused by
# spiders ballooning their way from one web to the other) - and these need to be removed. 

excess_spiders <- read.csv('excess_spiders.csv')
colonies_to_remove <- subset(excess_spiders, excess_spiders_observed==TRUE)$colonyname

#load function which removes all rows in the colonynames list/vector
source('remove_colonies.R')

wts <- remove_colonies(wts_raw, as.vector(colonies_to_remove),'colonyname')

wts<-subset(wts,wts$groupsize!='NA') # cleaning up the data frame once more to avoid an extra 'NA' group ! 
wts$wtloss<-wts$b4wt-wts$aftrwt

  
hqt<-round(quantile(wts$wtloss,0.975,na.rm=T),4)
lqt<-round(quantile(wts$wtloss,0.025,na.rm=T),4)
medn<-round(quantile(wts$wtloss,0.5,na.rm=T),4)


# now converting all the wts into milligrams with 1 decimal point accuracy - this is just to ease interpretation

wts$wtlossmg<-round(wts$wtloss*1000,1)
wts$b4wtmg<-round(wts$b4wt*1000,1)
wts$aftrwtmg<-round(wts$aftrwt*1000,1)
wts$wtlossmg<-round(wts$wtloss*1000,1)


# how much of missing data -(for Weightloss)  is there in this dataset ?
rws<-dim(wts)[1]
mssng<-sum(is.na(wts$wtloss))
pstagemiss<-mssng/rws

#are weight loss and spider size connected ? - let's check this out ! 
plot(wts$cephwid,wts$wtloss*1000,pch=20,col='red')
cor.test(wts$cephwid,wts$wtloss,method='spearman')

```

Now, let's build the models.

```{r,echo=TRUE}
library(lme4)

# both of these results suggest that there's an influence of previous weight, and size on silk investment
# maybe it's best to do some kind of mixed model to figure this out :

M0<-lm(wtlossmg~b4wtmg+wtday+groupsize+cephwid,data=wts)
summary(M0)
```
M0 suggests that groupsize makes a difference - but this is a very simple model 

```{r, echo=TRUE}
ml0<-lmer(wtlossmg~b4wtmg+wtday+groupsize+cephwid+(1|colonyname),data=wts)
summary(ml0)
#I want to check if 'spider ID' itself has an effect on wt loss !!! 
# now I've to created a combination of two columns to assign each spider a completely unique id 
# across the experiment.
wts$uniqid<-paste(wts$colonyname,wts$ID,sep='')

ml1<-lmer(wtlossmg~b4wtmg+wtday+groupsize+cephwid+(1|colonyname)+(1|uniqid),data=wts)
summary(ml1)
```

This shows that there's no effect of colonyname as such  on the amount of weight loss, but rather that smaller sized spiders tend  to invest more silk, and being in a good condition leads to more silk investment. There probably isn't enough sample size to actually estimate the variation caused by individuals - and so this analysis is inconclusive. 

Let's test for the effect of batch - this might affect the fact that
# different batches were run at different points of time.
```{r, echo=TRUE}
lastchar<-function(V){substr(V,nchar(V),nchar(V))}

wts$batchid<-lastchar(as.character(wts$colonyname))

ml0_2<-lmer(wtlossmg~b4wtmg+wtday+groupsize+cephwid+(1|colonyname)+(1|batchid),data=wts)

summary(ml0_2)
```

...and so - batch ID doesn't have an effect on the weight loss - this is good to know!

#diagnostic plots for the ml0_2 :
```{r,echo=TRUE}
P0<-hist(resid(ml0_2))
P0
P1<-plot(resid(ml0_2))
P1
# the residual analysis looks pretty decent.

# - just checking if this is caused by some data :

#choosing a subset of cephwidths
wsbt<-subset(wts,cephwid>1.9 & cephwid<2.8)

ml0_2_testcephwids<-lmer(wtlossmg~b4wtmg+wtday+groupsize+cephwid+(1|colonyname)+(1|batchid),data=wsbt)
summary(ml0_2)
# --the effects still remain the same

# choosing a subset of weightloss-es

wtlsbt<-subset(wts,wtlossmg<30 & wtlossmg>1)

ml0_2_testwts<-lmer(wtlossmg~b4wtmg+wtday+groupsize+cephwid+(1|colonyname)+(1|batchid),data=wtlsbt)
plot(resid(ml0_2_testwts))

# is it the effect of the error in rouding off of the cephwid?

wtscephtst<-wts;
wtscephtst$cephwidrnd<-round(wtscephtst$cephwid,1)
ml0_2_testceph<-lmer(wtlossmg~b4wtmg+wtday+groupsize+cephwidrnd+(1|colonyname)+(1|batchid),data=wtscephtst)

# the effect of mimicking measurement error still remain the same 
wts$bodycondn<-wts$b4wtmg/wts$cephwid

ml0_2_bodycndn<-lmer(wtlossmg~bodycondn+wtday+groupsize+(1|colonyname)+(1|batchid),data=wts)

# anyway you can still calculated the %age of spiders that were re-weighed:

# number of spiders weighed
numwd<-sum(is.na(wts$b4wt)==F)
# number of spiders re-weighed
numrewd<-sum(is.na(wts$aftrwt)==F)

# percentage of spiders re-weighed: ~80%
pctgrwd<-numrewd/numwd

# -- average weight of spiders b4 :
mean(wts$b4wtmg,na.rm=T)
```


### Other descriptive data

#### Cephalothorax widths

```{r, echo=TRUE}
cephalothorax_width = wts$cephwid
mean_cephwid <- mean(cephalothorax_width, na.rm=TRUE)
std_cephwid <- sd(cephalothorax_width, na.rm=TRUE)
sample_size = length(cephalothorax_width) - length(sum(is.na(cephalothorax_width)))
sprintf('Mean cephalothorax width +/- sd: %f +/- %f mm, N=%d', mean_cephwid, std_cephwid, sample_size)
```


```{r, echo=TRUE}
mean_startwt <- mean(wts$b4wt, na.rm=TRUE)
std_startwt <- sd(wts$b4wt, na.rm=TRUE)
sample_size = length(wts$b4wt) - length(sum(is.na(wts$b4wt)))

sprintf('Mean cephalothorax width +/- sd: %f +/- %f mg, N=%d', mean_startwt*1000, std_startwt*1000, sample_size)

```
